---
title: "`RobustGaSP`: MLE vs MAP estimation"
author: "37D real data"
date: "AOD vs AOD"
output: 
  pdf_document:
      number_sections: true
fontsize: 10pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{mybgcolor}{HTML}{ffebf5}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RobustGaSP)
# library(DiceKriging)
library(knitr)
library(readr)
library(lhs)
library(tidyverse)
library(ncdf4)
library(fields)
library(maps)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html")),list.files(pattern="*archived")),list.files(pattern="*aux")),list.files(pattern="GAM_analysis.R"))
                  )
     ) 
  {
  source(f)
}
rm(f, "loo_validation_summary", "m_pred_plots", "prop_wn_95CI", "MSE")
options(digits = 22)
```

<!-- \pagecolor{mybgcolor} -->

```{r, echo=F}
input <- 
  as.matrix(read.csv("X:/johnson_group/Aerosol-MFR/A-CURE-UKESM1-PPE-Data/Design/ACURE-UKESM1-PPE-par-norm-Design.csv"))

# dimensional of the inputs
dim_inputs <- ncol(input)
# number of the inputs
num_obs <- nrow(input)
# starting point for data
start_point <- 0
# input names
input_names <- colnames(input)
```

```{r, echo=F, message=F}
## We import the NetCDF file, then extract longitudes and
## latitudes, adjusting the longitudes ready for producing maps
response_AOD_nc <- nc_open(
paste("X:/johnson_group/Aerosol-MFR/A-CURE-UKESM1-PPE-Data/Outputs/AOD/ACURE_P3_AOD_Total_jul.nc")
)
longitude = ncvar_get(response_AOD_nc,"longitude") - 180
latitude = ncvar_get(response_AOD_nc,"latitude")

## Remove all unneeded objects
rm("response_AOD_nc")
```

```{r, echo=F}
long <- 5; lat <- 114
```

# Variable 1: AOD at longitude `r longitude[long]` and latitude `r latitude[lat]`

We start with AOD at the gridbox highlighted in green in Figure 1.

```{r, fig.cap="Selected gridbox is highlighted in green", echo=F}
map_matrix <- matrix(NA, nrow = 192, ncol = 144)
map_matrix[5:6, 114] <- 0
map_matrix[168:169, 102] <- 0.5
map_matrix[long, lat] <- 0.9

image.plot(longitude, latitude, 
           rbind(map_matrix[97:192,],
                 map_matrix[1:96,]),
           breaks = c(0,0.1,0.2,0.3,0.4,0.5,
                      0.6,0.7,0.8,0.9,1),
           col = c("blue","red3","red1","indianred1", "magenta",
                   "honeydew1", "darkseagreen1", "palegreen2", "seagreen3", 
                   "springgreen4"),
           xlab = "longitude", ylab = "latitude",
           axis.args=list(
             at=c(0,0.1,0.2,0.3,0.4,0.5,
                  0.6,0.7,0.8,0.9,1),labels=as.character(c(0,0.1,0.2,0.3,0.4,0.5,
                                                           0.6,0.7,0.8,0.9,1)),mgp=c(3,0.5,0)
           )
)
maps::map("world",lwd=1.2,add=TRUE, lty=1, col = "black")
```

## Emulator fit and validation

```{r, echo=F}
response_AOD_nc <- nc_open(
paste("X:/johnson_group/Aerosol-MFR/A-CURE-UKESM1-PPE-Data/Outputs/AOD/ACURE_P3_AOD_Total_jul.nc")
)
response_AOD_all_levels_array <- ncvar_get(response_AOD_nc, names(response_AOD_nc$var))

## We only want level 2, so extract that as a 192 × 144 × 221 array
response_AOD_array <- response_AOD_all_levels_array[,,2,]

## Remove all unneeded objects
rm("response_AOD_nc", "response_AOD_all_levels_array")
```

```{r, echo=F}
output <- matrix(response_AOD_array[long,lat,], nrow = num_obs)
```

SA gets done.

```{r, include=F}
input_names[7]
hist(input[,7], xlim=c(0,1))
```


```{r, echo=F}
# sum(input[,7] >= 0.1)
prim_so4_diam_low_indices <- which(input[,7] >= 0.1)
input <- input[prim_so4_diam_low_indices,]
hist(input[,7], xlim=c(0,1))
num_obs <- nrow(input)
num_obs

# includedYN_AODjuly_inputs <- data.frame(cbind(input[,7] >= 0.1, output, input))
# colnames(includedYN_AODjuly_inputs) <- c("included", "AOD_july", input_names)
# write.csv(includedYN_AODjuly_inputs, "C:/Users/smp22ijw/Downloads/includedYN_AODjuly_inputs.csv", row.names = F)
```

```{r, echo=F}
output <- matrix(response_AOD_array[long,lat,prim_so4_diam_low_indices], nrow = num_obs)
range(output)
```

```{r, echo=F}
colnames(output) <- "AOD"
input_output <- cbind(input, output)

## Plots
# input_output_ordered <- input_output %>% 
#   as_tibble %>%
#   gather('parameter', 'value', -AOD)
# input_output_ordered <- transform(input_output_ordered,
#                        parameter=factor(parameter, levels = c(input_names)))
# input_output_ordered %>%
#   ggplot(aes(x=value, y=AOD)) +
#   geom_point(alpha = 0.5) +
#   facet_wrap(~parameter) +
#   labs(y='output', x='input') +
#   geom_smooth(method = mgcv::gam,
#               formula = y ~ s(x, bs = "tp"),
#               fill = "red",
#               method.args = list(method="GCV.Cp"))
mainEffects <- rep(0, 37)
for(i in 1:37){
  gam <- mgcv::gam(input_output[, 38] ~ s(input_output[, i]))
  mainEffects[i] <- var(gam$fitted) / var(input_output[, 38])
}
# round(mainEffects * 100)
# Increase margin size
# par(mar=c(1,3,1,0.5))
# par(mar=c(6.5,3,1,0.5))
barplot(mainEffects,
        # names.arg = "",
        names.arg = c(input_names),
        las=2, cex.names=0.75,
        ylim = c(0,0.7)
)
th <- 0.03
input_names[mainEffects >= th]
mainEffects[mainEffects >= th]
sum(mainEffects[mainEffects >= th])

```

```{r}
# input <- input[,which(mainEffects >= th)]
```

```{r}
ncol(input)
```


We have a `r dim_inputs`-dimensional input, one output (AOD in July at the gridbox with longitude `r longitude[long]` and latitude `r latitude[lat]`), and `r num_obs` noise-free observations. We will assume a linear mean function of the form

$\hspace{2.6cm}  m(x) = \beta_0 + \beta_1$(`r input_names[1]`) $x_1 + \beta_2$(`r input_names[2]`) $+ \dots + \beta_{37}$(`r input_names[37]`)

The Gauss (squared exponential) *and* Matern 5/2 kernels will be compared for both estimation methods.

```{r, echo=FALSE}
H <- H_maker(input)
```

```{r, include=F}
m.mle.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     method = "mle",                                              # maximum likelihood estimation
                     nugget.est=TRUE
                     )

m.mle.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     method = "mle",                                              # maximum likelihood estimation
                     nugget.est=TRUE
                     )

m.map.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     nugget.est=TRUE
                     )

m.map.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     nugget.est=TRUE
                     )
```

```{r, echo=F}
# number of points to be predicted 
num_testing_input <- 100

# generate points to be predicted
# testing_input <- maximinLHS(n=num_testing_input, k=dim_inputs)

## Export 
# write_lines(as.vector(testing_input),
#             file=
#               paste0(
#                 "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_",
#                 dim_inputs,
#                 "D_", num_testing_input, "newpts.txt")
#             )

## Import
testing_input <-
  matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_", dim_inputs, "D_", num_testing_input, "newpts.txt"))),
         ncol = dim_inputs)
testing_input[,7] <- 0.1 + testing_input[,7] * 0.9
dim_inputs <- ncol(input)
# testing_input <- testing_input[,which(mainEffects >= th)]
```

```{r, echo=F}
H_testing <- H_maker(testing_input)
```

```{r, echo=F}
m.mle.gauss.predict <- RobustGaSP::predict(m.mle.gauss, testing_input, testing_trend = H_testing)

m.mle.matern52.predict <- RobustGaSP::predict(m.mle.matern52, testing_input, testing_trend = H_testing)

m.map.gauss.predict <- RobustGaSP::predict(m.map.gauss, testing_input, testing_trend = H_testing)

m.map.matern52.predict <- RobustGaSP::predict(m.map.matern52, testing_input, testing_trend = H_testing)
```

`r num_testing_input` new LH points are generated and predictions at these points made. The mean width of the 95% CI's for these predictions are shown in Table 1.

```{r, echo=F}
# average length of posterior predictive credible interval

length_95CI_mle_gauss <- length_95CI(m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95)

length_95CI_mle_matern52 <- length_95CI(m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95)

length_95CI_map_gauss <- length_95CI(m.map.gauss.predict$lower95, m.map.gauss.predict$upper95)

length_95CI_map_matern52 <- length_95CI(m.map.matern52.predict$lower95, m.map.matern52.predict$upper95)
```

```{r, echo=F}
checks_per_estimation <- data.frame(
  cbind(
    c(rep("MLE",2), rep("MAP",2)),
    rep(c("Gauss", "Matern 5/2"), 2),
    round(c(length_95CI_mle_gauss, length_95CI_mle_matern52, length_95CI_map_gauss, length_95CI_map_matern52),4)
    )
  )
colnames(checks_per_estimation) <- c("Estimate method", "Kernel", "Mean width of 95% CI")
kable(checks_per_estimation, 
      align=c("l","l","c"),
      caption = "Mean width of 95% CI"
      )
```

Leave-one-out validation is carried out; 
<!-- for each of the observations, fitting a GP using all observations except this one, then using the fitted emulator to predict the output at the excluded observation's input setting. The 95% credible interval for each prediction is checked against the actual output value, and  -->
the results are shown in Figure 2.

```{r, eval=F, echo=F}
m_mle_loo(input, output, "gauss")
m_mle_loo(input, output, "matern52")
m_map_loo(input, output, "gauss")
m_map_loo(input, output, "matern52")
```

```{r, eval=F, fig.show="hold", out.width="50%", fig.height=5, fig.cap="LOO CV. Top: MLE with Gauss (left) and Matern 5/2 (right) kernels. Bottom: MAP.", eval=F, echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
loo_validation_plot(m.mle.gauss.loo)

loo_validation_plot(m.mle.matern52.loo)

# MAP plots
loo_validation_plot(m.map.gauss.loo)

loo_validation_plot(m.map.matern52.loo)

# MLE proportion validated
round(sum(m.mle.gauss.loo$validated) / num_obs, 2)

round(sum(m.mle.matern52.loo$validated) / num_obs, 2)

# MAP proportion validated
round(sum(m.map.gauss.loo$validated) / num_obs, 2)

round(sum(m.map.matern52.loo$validated) / num_obs, 2)

```

Going forward, for now, only the Gauss kernel is dealt with; Matern 5/2 will be returned to once I've figured out the estimates returned by the two packages and the derivative of the Matern 5/2 kernel. 

```{r, echo=F}
# parameter estimates

m_ests(m.mle.gauss, "mle", "gauss")

m_ests(m.map.gauss, "map", "gauss")
```

## Verifying the estimates by-hand for both estimation methods

To calculate the posterior predictions at the `r num_testing_input` testing points by-hand, we require the matrix $H$ containing the inputs for the `r num_obs` observations and well as matrix $A$, its inverse, and $\textbf{t}(x^{*})$ for all `r num_obs` of the new inputs $x^{*}$. $A$ and $\textbf{t}(x^{*})$ are built using the estimate of the range parameter returned by each package. By-hand predictions match those returned by the two estimation methods (Figure 3).

Note that for `solve`, which is a numerical way to calculate the inverse of a matrix, if some of the numbers used in the inverse calculation for the matrix `A` are very small, it assumes that they are zero, leading to the assumption that it is a singular matrix. This is why it specifies that they are computationally singular, because the matrix might not be singular given a different tolerance. The solution is to set the tolerance for when it assumes that they are zero. `solve` allows you to specify this parameter (`tol`).

```{r, echo=F}
# make A and A_inv

# using range estimate(s) from MLE
m.mle.gauss.A <- A_maker(input, m.mle.gauss.ranges)
m.mle.gauss.A <- m.mle.gauss.A + diag(m.mle.gauss@nugget, num_obs)
m.mle.gauss.A.inv <- solve(m.mle.gauss.A)

# using range estimate(s) from MAP
m.map.gauss.A <- A_maker(input, m.map.gauss.ranges)
m.map.gauss.A <- m.map.gauss.A + diag(m.map.gauss@nugget, num_obs)
m.map.gauss.A.inv <- solve(m.map.gauss.A)

# make t

# using MLE range estimate(s)
m.mle.gauss.t <- t_maker(input, testing_input, m.mle.gauss.ranges)

# using MAP range estimate(s)
m.map.gauss.t <- t_maker(input, testing_input, m.map.gauss.ranges)
```

```{r, fig.show="hold", out.width="50%", fig.cap="The predictions found by-hand match well with those made by both estimation methods (left), with the latter minus the former found to be essentially zero (right). Top: MLE. Bottom: MAP.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
m.mle.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.mle.gauss.trends,
                                   m.mle.gauss.t,
                                   m.mle.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.mle.gauss.predict, m.mle.gauss.predictbyhand)

# MAP plots
m.map.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.map.gauss.trends,
                                   m.map.gauss.t,
                                   m.map.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.map.gauss.predict, m.map.gauss.predictbyhand)

plot(m.mle.gauss.predict$mean ~ m.map.gauss.predict$mean)

plot(m.mle.gauss.predictbyhand ~ m.map.gauss.predictbyhand)

```

```{r, echo = F}

m.mle.gauss.score <- sum(abs(m.mle.gauss.predict$mean - as.vector(m.mle.gauss.predictbyhand))) / (10^(-8)*num_testing_input)

m.map.gauss.score <- sum(abs(m.map.gauss.predict$mean - as.vector(m.map.gauss.predictbyhand))) / (10^(-8)*num_testing_input)

m.mle.gauss.score
m.map.gauss.score
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{N}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{N},$$ where $N$ is the number of test points, then divide this by $10^{-8}$ (machine precision) to get `r signif(m.mle.gauss.score, 1)` for MLE and `r signif(m.map.gauss.score, 1)` for MAP.

```{r, echo=F}
m.mle.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.mle.gauss.pds[,r] <-
    matrix(m.mle.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.mle.gauss.ranges[r]^2)) *
       m.mle.gauss.t) %*% 
    m.mle.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.mle.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.mle.gauss.pdsnormed.1 <- m.mle.gauss.pds / sqrt(rowSums(m.mle.gauss.pds^2))

m.map.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.map.gauss.pds[,r] <-
    matrix(m.map.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.map.gauss.ranges[r]^2)) *
       m.map.gauss.t) %*% 
    m.map.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.map.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.map.gauss.pdsnormed.1 <- m.map.gauss.pds / sqrt(rowSums(m.map.gauss.pds^2))
```

We create an object containing the emulated partial derivatives for use in the alignment measure. Finally, we create an object containing the normalised partial derivatives for use in the alignment measure.

```{r, echo=F}
rm(list=setdiff(ls(), 
                c("actual.1", "dim_inputs", "H", "H_testing", "input", "input_names", "longitude", "latitude", "m.mle.gauss.pdsnormed.1", "m.map.gauss.pdsnormed.1", "num_obs", "num_testing_input", "output", "prim_so4_diam_low_indices", "response_AOD_array", "testing_input", "testing_output")))

library(RobustGaSP)
# library(DiceKriging)
library(knitr)
library(readr)
library(lhs)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html")),list.files(pattern="*archived")),list.files(pattern="*aux")),list.files(pattern="GAM_analysis.R"))
                  )
     ) 
  {
  source(f)
}
rm(f, "loo_validation_summary", "m_pred_plots", "prop_wn_95CI", "MSE")
```

```{r, echo=F}
long <- 6; lat <- 114
```

# Variable 2: AOD at longitude `r longitude[long]` and latitude `r latitude[lat]`

We start with AOD at the gridbox highlighted in green in Figure 1.

```{r, fig.cap="Selected gridbox is highlighted in green", echo=F}
map_matrix <- matrix(NA, nrow = 192, ncol = 144)
map_matrix[5:6, 114] <- 0
map_matrix[168:169, 102] <- 0.5
map_matrix[long, lat] <- 0.9

image.plot(longitude, latitude, 
           rbind(map_matrix[97:192,],
                 map_matrix[1:96,]),
           breaks = c(0,0.1,0.2,0.3,0.4,0.5,
                      0.6,0.7,0.8,0.9,1),
           col = c("blue","red3","red1","indianred1", "magenta",
                   "honeydew1", "darkseagreen1", "palegreen2", "seagreen3", 
                   "springgreen4"),
           xlab = "longitude", ylab = "latitude",
           axis.args=list(
             at=c(0,0.1,0.2,0.3,0.4,0.5,
                  0.6,0.7,0.8,0.9,1),labels=as.character(c(0,0.1,0.2,0.3,0.4,0.5,
                                                           0.6,0.7,0.8,0.9,1)),mgp=c(3,0.5,0)
           )
)
maps::map("world",lwd=1.2,add=TRUE, lty=1, col = "black")
```

## Emulator fit and validation

```{r, echo=F}
# output <- matrix(response_AOD_array[long,lat,], nrow = num_obs)
output <- matrix(response_AOD_array[long,lat,prim_so4_diam_low_indices], nrow = num_obs)
range(output)
```

We have a `r dim_inputs`-dimensional input, one output (AOD in July at the gridbox with longitude `r longitude[long]` and latitude `r latitude[lat]`), and `r num_obs` noise-free observations. We will assume a linear mean function of the form

$\hspace{2.6cm}  m(x) = \beta_0 + \beta_1$(`r input_names[1]`) $x_1 + \beta_2$(`r input_names[2]`) $+ \dots + \beta_{37}$(`r input_names[37]`)

The Gauss (squared exponential) *and* Matern 5/2 kernels will be compared for both estimation methods.

```{r, include=F}
m.mle.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     method = "mle",                                              # maximum likelihood estimation
                     nugget.est=TRUE
                     )

m.mle.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     method = "mle",                                              # maximum likelihood estimation
                     nugget.est=TRUE
                     )

m.map.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     nugget.est=TRUE
                     )

m.map.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     nugget.est=TRUE
                     )
```

```{r perform-predictions, echo=F}
m.mle.gauss.predict <- RobustGaSP::predict(m.mle.gauss, testing_input, testing_trend = H_testing)

m.mle.matern52.predict <- RobustGaSP::predict(m.mle.matern52, testing_input, testing_trend = H_testing)

m.map.gauss.predict <- RobustGaSP::predict(m.map.gauss, testing_input, testing_trend = H_testing)

m.map.matern52.predict <- RobustGaSP::predict(m.map.matern52, testing_input, testing_trend = H_testing)
```

`r num_testing_input` new LH points are generated and predictions at these points made. The mean width of the 95% CI's for these predictions are shown in Table 1.

```{r, echo=F}
# average length of posterior predictive credible interval

length_95CI_map_gauss <- length_95CI(m.map.gauss.predict$lower95, m.map.gauss.predict$upper95)

length_95CI_map_matern52 <- length_95CI(m.map.matern52.predict$lower95, m.map.matern52.predict$upper95)

length_95CI_mle_gauss <- length_95CI(m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95)

length_95CI_mle_matern52 <- length_95CI(m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95)
```

```{r, echo=F}
checks_per_estimation <- data.frame(
  cbind(
    c(rep("MLE",2), rep("MAP",2)),
    rep(c("Gauss", "Matern 5/2"), 2),
    round(c(length_95CI_mle_gauss, length_95CI_mle_matern52, length_95CI_map_gauss, length_95CI_map_matern52),4)
    )
  )
colnames(checks_per_estimation) <- c("Estimate method", "Kernel", "Mean width of 95% CI")
kable(checks_per_estimation, 
      align=c("l","l","c"),
      caption = "Mean width of 95% CI"
      )
```

Leave-one-out validation is carried out; 
<!-- for each of the observations, fitting a GP using all observations except this one, then using the fitted emulator to predict the output at the excluded observation's input setting. The 95% credible interval for each prediction is checked against the actual output value, and  -->
the results are shown in Figure 2.

```{r, eval=F, echo=F}
m_mle_loo(input, output, "gauss")
m_mle_loo(input, output, "matern52")
m_map_loo(input, output, "gauss")
m_map_loo(input, output, "matern52")
```

```{r, eval=F, fig.show="hold", out.width="50%", fig.height=5, fig.cap="LOO CV. Top: MLE with Gauss (left) and Matern 5/2 (right) kernels. Bottom: MAP.", eval=F, echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
loo_validation_plot(m.mle.gauss.loo)

loo_validation_plot(m.mle.matern52.loo)

# MAP plots
loo_validation_plot(m.map.gauss.loo)

loo_validation_plot(m.map.matern52.loo)

```

Going forward, for now, only the Gauss kernel is dealt with; Matern 5/2 will be returned to once I've figured out the estimates returned by the two packages and the derivative of the Matern 5/2 kernel. 

```{r, echo=F}
# parameter estimates

m_ests(m.mle.gauss, "mle", "gauss")

m_ests(m.map.gauss, "map", "gauss")
```

## Verifying the estimates by-hand for both estimation methods

To calculate the posterior predictions at the `r num_testing_input` testing points by-hand, we require the matrix $H$ containing the inputs for the `r num_obs` observations and well as matrix $A$, its inverse, and $\textbf{t}(x^{*})$ for all `r num_obs` of the new inputs $x^{*}$. $A$ and $\textbf{t}(x^{*})$ are built using the estimate of the range parameter returned by each package. By-hand predictions match those returned by the two estimation methods (Figure 3).

Note that for `solve`, which is a numerical way to calculate the inverse of a matrix, if some of the numbers used in the inverse calculation for the matrix `A` are very small, it assumes that they are zero, leading to the assumption that it is a singular matrix. This is why it specifies that they are computationally singular, because the matrix might not be singular given a different tolerance. The solution is to set the tolerance for when it assumes that they are zero. `solve` allows you to specify this parameter (`tol`).

```{r, echo=F}
# make A and A_inv

# using range estimate(s) from MLE
m.mle.gauss.A <- A_maker(input, m.mle.gauss.ranges)
m.mle.gauss.A <- m.mle.gauss.A + diag(m.mle.gauss@nugget, num_obs)
m.mle.gauss.A.inv <- solve(m.mle.gauss.A)

# using range estimate(s) from MAP
m.map.gauss.A <- A_maker(input, m.map.gauss.ranges)
m.map.gauss.A <- m.map.gauss.A + diag(m.map.gauss@nugget, num_obs)
m.map.gauss.A.inv <- solve(m.map.gauss.A)

# make t

# using MLE range estimate(s)
m.mle.gauss.t <- t_maker(input, testing_input, m.mle.gauss.ranges)

# using MAP range estimate(s)
m.map.gauss.t <- t_maker(input, testing_input, m.map.gauss.ranges)
```

```{r, fig.show="hold", out.width="50%", fig.cap="The predictions found by-hand match well with those made by both estimation methods (left), with the latter minus the former found to be essentially zero (right). Top: MLE. Bottom: MAP.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
m.mle.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.mle.gauss.trends,
                                   m.mle.gauss.t,
                                   m.mle.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.mle.gauss.predict, m.mle.gauss.predictbyhand)

# MAP plots
m.map.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.map.gauss.trends,
                                   m.map.gauss.t,
                                   m.map.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.map.gauss.predict, m.map.gauss.predictbyhand)

plot(m.mle.gauss.predict$mean ~ m.map.gauss.predict$mean)

plot(m.mle.gauss.predictbyhand ~ m.map.gauss.predictbyhand)

```

```{r, echo = F}

m.mle.gauss.score <- sum(abs(m.mle.gauss.predict$mean - as.vector(m.mle.gauss.predictbyhand))) / (10^(-8)*num_testing_input)

m.map.gauss.score <- sum(abs(m.map.gauss.predict$mean - as.vector(m.map.gauss.predictbyhand))) / (10^(-8)*num_testing_input)
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{N}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{N},$$ where $N$ is the number of test points, then divide this by $10^{-8}$ (machine precision) to get `r signif(m.mle.gauss.score, 1)` for MLE and `r signif(m.map.gauss.score, 1)` for MAP.

```{r, echo=F}
m.mle.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.mle.gauss.pds[,r] <-
    matrix(m.mle.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.mle.gauss.ranges[r]^2)) *
       m.mle.gauss.t) %*% 
    m.mle.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.mle.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.mle.gauss.pdsnormed.2 <- m.mle.gauss.pds / sqrt(rowSums(m.mle.gauss.pds^2))

m.map.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.map.gauss.pds[,r] <-
    matrix(m.map.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.map.gauss.ranges[r]^2)) *
       m.map.gauss.t) %*% 
    m.map.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.map.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.map.gauss.pdsnormed.2 <- m.map.gauss.pds / sqrt(rowSums(m.map.gauss.pds^2))
```

We create an object containing the emulated partial derivatives for use in the alignment measure. Finally, we create an object containing the normalised partial derivatives for use in the alignment measure.

# Alignment measure

```{r, echo=F}
AM.mle.1.2 <- mean(abs(rowSums(m.mle.gauss.pdsnormed.1 * m.mle.gauss.pdsnormed.2)))
AM.mle.1.2
AM.map.1.2 <- mean(abs(rowSums(m.map.gauss.pdsnormed.1 * m.map.gauss.pdsnormed.2)))
AM.map.1.2
```