---
title: "`RobustGaSP`: MLE vs MAP estimation"
author: "2D test example"
date: "$y = \\sqrt{2 - x_1^2 - x_2^2}$ vs Branin-Hoo"
output: 
  pdf_document:
      number_sections: true
fontsize: 10pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{mybgcolor}{HTML}{ffebf5}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RobustGaSP)
# library(DiceKriging)
library(knitr)
library(readr)
library(lhs)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html")),list.files(pattern="*archived"))
                  )
     ) 
  {
  source(f)
  }
```

<!-- \pagecolor{mybgcolor} -->

```{r, echo=F}
# dimensional of the inputs
dim_inputs <- 2   
# number of the inputs
num_obs <- 10*dim_inputs
# starting point for data
start_point <- 0
# lhs design
# input <- maximinLHS(n=num_obs, k=dim_inputs)  ##maximin lhd sample
## Export 
# write_lines(as.vector(input),
#             file=
#               paste0(
#                 "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/input_",
#                 dim_inputs,
#                 "D_",
#                 num_obs,
#                 "obs.txt")
#             )

## Import
input <-
  matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/input_", dim_inputs, "D_", num_obs, "obs.txt"))),
         ncol = dim_inputs)
```

# Function 1: $y = \sqrt{2 - x_1^2 - x_2^2}$

## Emulator fit and validation

```{r, echo=F}

# outputs
sphere_eighth_bleft <- function(inputs) {
  sqrt(2 - inputs[1]^2 - inputs[2]^2)
  }

output <- matrix(apply(input, 1, sphere_eighth_bleft), nrow = num_obs)

```

We have a `r dim_inputs`-dimensional input, one output, and `r num_obs` noise-free observations (Figure 1).

```{r, out.width="45%", fig.height = 6.75, echo=F, fig.cap="Contour plot for the function $y = \\sqrt{2 - x_1^2 - x_2^2}$, with the location in the 2D input space of the observations' input settings shown in blue.", fig.align = 'center', warning=F}

x <- seq(-2,2, by=0.1)
y <- seq(-2,2, by=0.1)
z_f <- sqrt(2^2 - outer(x ^ 2, y ^ 2, "+"))

contour(x[21:31], y[21:31], z_f[21:31, 21:31], xlim = c(0,1), ylim = c(0,1), nlevels = 40, labcex=1, xlab="x1", ylab="x2")
points(input, add=T, pch=19, col="blue")
```

Going forward, we will assume a linear mean function of the form $m(x) = \beta_0 + \beta_1x_1 + \beta_2x_2.$ The Gauss (squared exponential) *and* Matern 5/2 kernels will be compared for both estimation methods.

```{r, echo=FALSE}
H <- H_maker(input)
```

```{r, include=F}
m.mle.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     method = "mle"                                               # maximum likelihood estimation
                     )

m.mle.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     method = "mle"                                               # maximum likelihood estimation
                     )

m.map.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     )

m.map.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No"
                     )
```

```{r testing-inputs, echo=F}
# number of points to be predicted 
num_testing_input <- 100

# generate points to be predicted
# testing_input <- maximinLHS(n=num_testing_input, k=dim_inputs)

## Export 
# write_lines(as.vector(testing_input),
#             file=
#               paste0(
#                 "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_",
#                 dim_inputs,
#                 "D_", num_testing_input, "newpts.txt")
#             )

## Import
testing_input <-
  matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_", dim_inputs, "D_", num_testing_input, "newpts.txt"))),
         ncol = dim_inputs)
```

```{r, echo=F}
H_testing <- H_maker(testing_input)
```

```{r perform-predictions, echo=F}
m.mle.gauss.predict <- RobustGaSP::predict(m.mle.gauss, testing_input, testing_trend = H_testing)

m.mle.matern52.predict <- RobustGaSP::predict(m.mle.matern52, testing_input, testing_trend = H_testing)

m.map.gauss.predict <- RobustGaSP::predict(m.map.gauss, testing_input, testing_trend = H_testing)

m.map.matern52.predict <- RobustGaSP::predict(m.map.matern52, testing_input, testing_trend = H_testing)
```

`r num_testing_input` new LH points are generated, and out-of-sample checks (as per `RobustGaSP`'s Help page) shown in Table 1, with the emulator predictions plotted against the actual outputs for these points shown in Figure 2.

```{r, echo=F}
# actual output values
testing_output <- matrix(apply(testing_input, 1, sphere_eighth_bleft), nrow = num_testing_input)

```

```{r prop_in_cred_int, echo=F}
# proportion covered by 95% posterior predictive credible interval

# m.mle.gauss.95CIprop
prop_wn_95CI_mle_gauss <- prop_wn_95CI(
  m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95
  )

prop_wn_95CI_mle_matern52 <- prop_wn_95CI(
  m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95
  )

prop_wn_95CI_map_gauss <- prop_wn_95CI(
  m.map.gauss.predict$lower95, m.map.gauss.predict$upper95
  )

prop_wn_95CI_map_matern52 <- prop_wn_95CI(
  m.map.matern52.predict$lower95, m.map.matern52.predict$upper95
  )
```

```{r MSE, echo=F}
# out of sample MSE

MSE_emulator_mle_gauss <- 
  MSE(m.mle.gauss.predict$mean, testing_output)

MSE_emulator_mle_matern52 <- 
  MSE(m.mle.matern52.predict$mean, testing_output)

MSE_emulator_map_gauss <- 
  MSE(m.map.gauss.predict$mean, testing_output)

MSE_emulator_map_matern52 <- 
  MSE(m.map.matern52.predict$mean, testing_output)
```

```{r length_of_cred_int, echo=F}
# average length of posterior predictive credible interval

length_95CI_map_gauss <- length_95CI(m.map.gauss.predict$lower95, m.map.gauss.predict$upper95)

length_95CI_map_matern52 <- length_95CI(m.map.matern52.predict$lower95, m.map.matern52.predict$upper95)

length_95CI_mle_gauss <- length_95CI(m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95)

length_95CI_mle_matern52 <- length_95CI(m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95)
```

```{r checks-summary, echo=F}
checks_per_estimation <- data.frame(
  cbind(
    c(rep("MLE",2), rep("MAP",2)),
    rep(c("Gauss", "Matern 5/2"), 2),
    round(c(prop_wn_95CI_mle_gauss, prop_wn_95CI_mle_matern52, prop_wn_95CI_map_gauss, prop_wn_95CI_map_matern52),2),
    signif(c(MSE_emulator_mle_gauss, MSE_emulator_mle_matern52, MSE_emulator_map_gauss, MSE_emulator_map_matern52),1),
    round(c(length_95CI_mle_gauss, length_95CI_mle_matern52, length_95CI_map_gauss, length_95CI_map_matern52),4)
    )
  )
colnames(checks_per_estimation) <- c("Estimate method", "Kernel", "Prop. w/n 95% CI", "Out-of-sample MSE", "Mean width of 95% CI")
kable(checks_per_estimation, 
      align=c("l","l","c","c","c"),
      caption = "Out-of-sample checks"
      )
```

```{r, fig.show="hold", out.width="50%", fig.height=5, fig.cap="Emulator vs model for Gauss (left) and Matern 5/2 (right). Red: MLE. Blue: MAP", echo=F}

par(mar = c(4, 4, .1, .1))

plot(testing_output, m.mle.gauss.predict$mean,
     col="red", pch=19, cex=0.5,
     xlab = "Model", ylab = "Emulator")
points(testing_output, m.map.gauss.predict$mean,
       col="blue", pch=19, cex=0.5)
abline(0,1)

plot(testing_output, m.mle.matern52.predict$mean,
     col="red", pch=19, cex=0.5,
     xlab = "Model", ylab = "Emulator")
points(testing_output, m.map.matern52.predict$mean,
     col="blue", pch=19, cex=0.5)
abline(0,1)
```

Finally, in preparation for the real world data where we cannot perform out-of-sample validation, leave-one-out validation is carried out; 
<!-- for each of the observations, fitting a GP using all observations except this one, then using the fitted emulator to predict the output at the excluded observation's input setting. The 95% credible interval for each prediction is checked against the actual output value, and  -->
the results are shown in Figure 3.

```{r, include=F}
m_mle_loo(input, output, "gauss")
m_mle_loo(input, output, "matern52")
m_map_loo(input, output, "gauss")
m_map_loo(input, output, "matern52")
```

```{r, validation-plots, fig.show="hold", out.width="50%", fig.height=5, fig.cap="LOO CV. Top: MLE with Gauss (left) and Matern 5/2 (right) kernels. Bottom: MAP.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
loo_validation_plot(m.mle.gauss.loo)

loo_validation_plot(m.mle.matern52.loo)

# MAP plots
loo_validation_plot(m.map.gauss.loo)

loo_validation_plot(m.map.matern52.loo)

```

\newpage

Going forward, for now, only the Gauss kernel is dealt with; Matern 5/2 will be returned to once I've figured out the estimates returned by the two packages and the derivative of the Matern 5/2 kernel. The estimates returned by both estimation methods using the Gauss kernel are shown in Table 2.

```{r, echo=F}
# parameter estimates

m_ests(m.mle.gauss, "mle", "gauss")

m_ests(m.map.gauss, "map", "gauss")
```

```{r, echo=F}
estimates_per_estimation <- data.frame(
  cbind(
    c("MLE", "MAP"),
    rep("Gauss", 2),
    round(c(m.mle.gauss.trends[1], m.map.gauss.trends[1]),2),
    round(c(m.mle.gauss.trends[2], m.map.gauss.trends[2]),2),
    round(c(m.mle.gauss.trends[3], m.map.gauss.trends[3]),2),
    round(c(m.mle.gauss.ranges[1], m.map.gauss.ranges[1]),2),
    round(c(m.mle.gauss.ranges[2], m.map.gauss.ranges[2]),2),
    round(c(m.mle.gauss.var, m.map.gauss.var),2)
    )
  )
colnames(estimates_per_estimation) <- c("Estimation", "Kernel", "Trend 1", "Trend 2", "Trend 3", "Range 1", "Range 2", "Variance")
kable(estimates_per_estimation, 
      align=c("l","l","c","c","c","c","c","c"),
      caption = "Estimates table"
      )
```

## Verifying the estimates by-hand for both estimation methods

To calculate the posterior predictions at the `r num_testing_input` testing points by-hand, we require the matrix $H$ containing the inputs for the `r num_obs` observations and well as matrix $A$, its inverse, and $\textbf{t}(x^{*})$ for all `r num_obs` of the new inputs $x^{*}$. $A$ and $\textbf{t}(x^{*})$ are built using the estimate of the range parameter returned by each package. By-hand predictions match those returned by the two estimation methods (Figure 4).

```{r, echo=F}
# make A and A_inv

# using range estimate(s) from MLE
m.mle.gauss.A <- A_maker(input, m.mle.gauss.ranges)
m.mle.gauss.A.inv <- solve(m.mle.gauss.A)

# using range estimate(s) from MAP
m.map.gauss.A <- A_maker(input, m.map.gauss.ranges)
m.map.gauss.A.inv <- solve(m.map.gauss.A)

# make t

# using MLE range estimate(s)
m.mle.gauss.t <- t_maker(input, testing_input, m.mle.gauss.ranges)

# using MAP range estimate(s)
m.map.gauss.t <- t_maker(input, testing_input, m.map.gauss.ranges)
```

```{r, fig.show="hold", out.width="50%", fig.cap="The predictions found by-hand match well with those made by both estimation methods (left), with the latter minus the former found to be essentially zero (right). Top: MLE. Bottom: MAP.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
m.mle.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.mle.gauss.trends,
                                   m.mle.gauss.t,
                                   m.mle.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.mle.gauss.predict, m.mle.gauss.predictbyhand)

# MAP plots
m.map.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.map.gauss.trends,
                                   m.map.gauss.t,
                                   m.map.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.map.gauss.predict, m.map.gauss.predictbyhand)

```

```{r, echo = F}

m.mle.gauss.score <- sum(abs(m.mle.gauss.predict$mean - as.vector(m.mle.gauss.predictbyhand))) / (10^(-8)*num_testing_input)

m.map.gauss.score <- sum(abs(m.map.gauss.predict$mean - as.vector(m.map.gauss.predictbyhand))) / (10^(-8)*num_testing_input)
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{N}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{N},$$ where $N$ is the number of test points, then divide this by $10^{-8}$ (machine precision) to get `r signif(m.mle.gauss.score, 1)` for MLE and `r signif(m.map.gauss.score, 1)` for MAP.

```{r, echo=F}
m.mle.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.mle.gauss.pds[,r] <-
    matrix(m.mle.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.mle.gauss.ranges[r]^2)) *
       m.mle.gauss.t) %*% 
    m.mle.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.mle.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.mle.gauss.pdsnormed.1 <- m.mle.gauss.pds / sqrt(rowSums(m.mle.gauss.pds^2))

m.map.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.map.gauss.pds[,r] <-
    matrix(m.map.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.map.gauss.ranges[r]^2)) *
       m.map.gauss.t) %*% 
    m.map.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.map.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.map.gauss.pdsnormed.1 <- m.map.gauss.pds / sqrt(rowSums(m.map.gauss.pds^2))
```

We create an object containing the emulated partial derivatives for use in the alignment measure, and check that these approximately match the partial derivatives of the actual function (Figure 5). Also note that the two estimation methods return similar estimates for the partial derivatives (Figure 6).

```{r, fig.show="hold", out.width="50%", fig.cap="The estimated partial derivatives found by each estimation method (red dots, left MLE and right MAP) match the true partial derivatives (black) well.", echo=F}

par(mar = c(4, 4, .1, .1))

# actual pds
sphere_eighth_bleft_ddx1 <- function(inputs) {
  0.5 * (-2*inputs[1]) / sqrt(2 - inputs[1]^2 - inputs[2]^2)
}
sphere_eighth_bleft_ddx2 <- function(inputs) {
  0.5 * (-2*inputs[2]) / sqrt(2 - inputs[1]^2 - inputs[2]^2)
}
actual.1 <- data.frame(cbind(rep(NA, num_testing_input), rep(NA, num_testing_input)))
colnames(actual.1) <- c("mean1","mean2")
actual.1$mean1 <- matrix(apply(testing_input, 1, sphere_eighth_bleft_ddx1), nrow = num_testing_input)
actual.1$mean2 <- matrix(apply(testing_input, 1, sphere_eighth_bleft_ddx2), nrow = num_testing_input)

# MLE plots
plot(actual.1, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(actual.1$mean1, m.mle.gauss.pds$X1), 
            max(actual.1$mean1, m.mle.gauss.pds$X1)),
     ylim=c(min(actual.1$mean2, m.mle.gauss.pds$X2), 
            max(actual.1$mean2, m.mle.gauss.pds$X2)),
     pch=19, cex=0.75
     )
points(m.mle.gauss.pds, pch=19, cex=0.75, col="red")
segments(actual.1$mean1, actual.1$mean2,
         m.mle.gauss.pds$X1, m.mle.gauss.pds$X2
         )

# MAP plots
plot(actual.1, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(actual.1$mean1, m.map.gauss.pds$X1), 
            max(actual.1$mean1, m.map.gauss.pds$X1)),
     ylim=c(min(actual.1$mean2, m.map.gauss.pds$X2), 
            max(actual.1$mean2, m.map.gauss.pds$X2)),
     pch=19, cex=0.75
     )
points(m.map.gauss.pds, pch=19, cex=0.75, col="red")
segments(actual.1$mean1, actual.1$mean2,
         m.map.gauss.pds$X1, m.map.gauss.pds$X2
         )

```

```{r, out.width="50%", fig.align="center", fig.cap="The estimated partial derivatives returned by each estimation method are similar (blue for MLE and red for MAP).", echo=F}

par(mar = c(4, 4, .1, .1))

plot(m.mle.gauss.pds, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(m.map.gauss.pds$X1, m.map.gauss.pds$X1), 
            max(m.map.gauss.pds$X1, m.map.gauss.pds$X1)),
     ylim=c(min(m.map.gauss.pds$X2, m.map.gauss.pds$X2), 
            max(m.map.gauss.pds$X2, m.map.gauss.pds$X2)),
     pch=19, cex=0.75, col="blue"
     )
points(m.map.gauss.pds, pch=19, cex=0.75, col="red")
segments(m.mle.gauss.pds$X1, m.mle.gauss.pds$X2,
         m.map.gauss.pds$X1, m.map.gauss.pds$X2
         )

```
Finally, we create an object containing the normalised partial derivatives for use in the alignment measure.

```{r, echo=F}
rm(list=setdiff(ls(), 
                c("actual.1", "dim_inputs", "H", "H_testing", "input", "m.mle.gauss.pdsnormed.1", "m.map.gauss.pdsnormed.1", "num_obs", "num_testing_input", "output", "testing_input", "testing_output")))

library(RobustGaSP)
# library(DiceKriging)
library(knitr)
library(readr)
library(lhs)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html")),list.files(pattern="*archived"))
                  )
     ) 
  {
  source(f)
  }
```

# Function 2: Branin-Hoo (`DiceKriging` version)

We repeat the above steps for the Brian-Hoo function, in the form it appears in the `DiceKriging` package, $$y = \left(x_2 - \frac{5}{4\pi^2}x_1^2 + \frac{5}{\pi}x_1 - 6\right)^2 + 10\left(1 - \frac{1}{8\pi}\right)\cos(x_1) + 10$$ (see Figure 7). The same training and testing points are used as for the first function.

```{r 2nd-functions-work-0, out.width="45%", fig.height = 6.75, echo=F, fig.cap="Contour plot for the Branin-Hoo function, with the location in the 2D input space of the observations' input settings shown in blue.", fig.align = 'center', warning=F}

n.grid <- 20
x.grid <- y.grid <- seq(0,1,length=n.grid)

# # Branin-Hoo parameter values
# a <- 1; b <- 5.1/(4*pi^2); c <- 5/pi; r <- 6; s <- 10; t <- 1/(8*pi)
# # Branin-Hoo 
# branin_hoo <- function(X, Y) {
#   a * (((Y) - b * (X)^2 + c * (X) - r)^2 + s * (1-t) * cos((X)) + s)
#   }
# # scaled Branin-Hoo 
# scaled_branin_hoo <- function(X, Y) {
#   a/51.95 * (((15*Y) - b * (15*X-5)^2 + c * (15*X-5) - r)^2 + s * (1-t) * cos((15*X-5)) - 44.81)
# }
# DiceKriging's scaled Branin-Hoo 
dk_scaled_branin_hoo <- function (x) {
    x1 <- x[1] * 15 - 5
    x2 <- x[2] * 15
    (x2 - 5/(4 * pi^2) * (x1^2) + 5/pi * x1 - 6)^2 + 10 * (1 - 
        1/(8 * pi)) * cos(x1) + 10
}

z.grid <- matrix(NA, nrow = n.grid, ncol = n.grid)
for (i in 1:n.grid) {
  for (j in 1:n.grid) {
    x <- x.grid[i]
    y <- y.grid[j]
    z.grid[i,j] <- dk_scaled_branin_hoo(c(x, y))
  }
}

contour(x.grid, y.grid, z.grid, xlim = c(0,1), ylim = c(0,1), nlevels = 40, labcex=1, xlab="x1", ylab="x2")

points(input, add=T, pch=19, col="blue")
```

\newpage

## Emulator fit and validation

Out-of-sample check results are shown in Table 3 and Figure 8.

```{r 2nd-functions-work-1, include=F, echo=F}
# output of train points
output <- matrix(apply(input, 1, dk_scaled_branin_hoo), nrow = num_obs)

# emulator build
m.mle.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     method = "mle"                                               # maximum likelihood estimation
                     )
m.mle.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     method = "mle"                                               # maximum likelihood estimation
                     )
m.map.gauss <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No",
                     kernel_type = 'pow_exp', alpha = rep(2, ncol(input)),        # squared exponential
                     )
m.map.matern52 <- rgasp(design = input, response = output, trend = H, lower_bound=FALSE, zero.mean = "No"
                     )

# predictions
m.mle.gauss.predict <- RobustGaSP::predict(m.mle.gauss, testing_input, testing_trend = H_testing)
m.mle.matern52.predict <- RobustGaSP::predict(m.mle.matern52, testing_input, testing_trend = H_testing)
m.map.gauss.predict <- RobustGaSP::predict(m.map.gauss, testing_input, testing_trend = H_testing)
m.map.matern52.predict <- RobustGaSP::predict(m.map.matern52, testing_input, testing_trend = H_testing)

# output of test points
testing_output <- matrix(apply(testing_input, 1, dk_scaled_branin_hoo), nrow = num_testing_input)

# proportion covered by 95% posterior predictive credible interval
prop_wn_95CI_mle_gauss <- prop_wn_95CI(
  m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95
  )
prop_wn_95CI_mle_matern52 <- prop_wn_95CI(
  m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95
  )
prop_wn_95CI_map_gauss <- prop_wn_95CI(
  m.map.gauss.predict$lower95, m.map.gauss.predict$upper95
  )
prop_wn_95CI_map_matern52 <- prop_wn_95CI(
  m.map.matern52.predict$lower95, m.map.matern52.predict$upper95
  )

# out of sample MSE
MSE_emulator_mle_gauss <- 
  MSE(m.mle.gauss.predict$mean, testing_output)
MSE_emulator_mle_matern52 <- 
  MSE(m.mle.matern52.predict$mean, testing_output)
MSE_emulator_map_gauss <- 
  MSE(m.map.gauss.predict$mean, testing_output)
MSE_emulator_map_matern52 <- 
  MSE(m.map.matern52.predict$mean, testing_output)

# average length of posterior predictive credible interval
length_95CI_map_gauss <- 
  length_95CI(m.map.gauss.predict$lower95, m.map.gauss.predict$upper95)
length_95CI_map_matern52 <- 
  length_95CI(m.map.matern52.predict$lower95, m.map.matern52.predict$upper95)
length_95CI_mle_gauss <- 
  length_95CI(m.mle.gauss.predict$lower95, m.mle.gauss.predict$upper95)
length_95CI_mle_matern52 <- 
  length_95CI(m.mle.matern52.predict$lower95, m.mle.matern52.predict$upper95)

```

```{r checks-summary-2, echo=F}
checks_per_estimation <- data.frame(
  cbind(
    c(rep("Maximum likelihood",2), rep("MAP",2)),
    rep(c("Gauss", "Matern 5/2"), 2),
    round(c(prop_wn_95CI_mle_gauss, prop_wn_95CI_mle_matern52, prop_wn_95CI_map_gauss, prop_wn_95CI_map_matern52),2),
    signif(c(MSE_emulator_mle_gauss, MSE_emulator_mle_matern52, MSE_emulator_map_gauss, MSE_emulator_map_matern52),1),
    round(c(length_95CI_mle_gauss, length_95CI_mle_matern52, length_95CI_map_gauss, length_95CI_map_matern52),4)
    )
  )
colnames(checks_per_estimation) <- c("Estimate method", "Kernel", "Prop. w/n 95% CI", "Out-of-sample MSE", "Mean width of 95% CI")
kable(checks_per_estimation, 
      align=c("l","l","c","c","c"),
      caption = "Out-of-sample checks"
      )
```

```{r, fig.show="hold", out.width="50%", fig.cap="Emulator versus model for the new points. Left: Using a Gauss kernel. Right: Using a Matern 5/2 kernel. Red: MLE. Blue: MAP", echo=F}

par(mar = c(4, 4, .1, .1))

plot(testing_output, m.mle.gauss.predict$mean,
     col="red", pch=19, cex=0.5,
     xlab = "Model", ylab = "Emulator")
points(testing_output, m.map.gauss.predict$mean,
       col="blue", pch=19, cex=0.5)
abline(0,1)

plot(testing_output, m.mle.matern52.predict$mean,
     col="red", pch=19, cex=0.5,
     xlab = "Model", ylab = "Emulator")
points(testing_output, m.map.matern52.predict$mean,
     col="blue", pch=19, cex=0.5)
abline(0,1)
```

LOO CV results are shown in Figure 9.

```{r 2nd-functions-work-2, include=F, echo=F}
# LOO CV
m_mle_loo(input, output, "gauss")
m_mle_loo(input, output, "matern52")
m_map_loo(input, output, "gauss")
m_map_loo(input, output, "matern52")
```

```{r, validation-plots-2, fig.show="hold", out.width="50%", fig.cap="LOO CV. Top: Maximum likelihood estimation for the Gauss (left) and Matern 5/2 (right) kernels. Bottom: Marginal posterior mode estimation for the Gauss (left) and Matern 5/2 (right) kernels.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
loo_validation_plot(m.mle.gauss.loo)

loo_validation_plot(m.mle.matern52.loo)

# MAP plots
loo_validation_plot(m.map.gauss.loo)

loo_validation_plot(m.map.matern52.loo)

```

\newpage

Parameter estimates using the Gauss kernel are shown in Table 4.

```{r 2nd-functions-work-3, include=F, echo=F}
# parameter estimates
m_ests(m.mle.gauss, "mle", "gauss")
m_ests(m.map.gauss, "map", "gauss")
```

```{r, echo=F}
estimates_per_estimation <- data.frame(
  cbind(
    c("MLE", "MAP"),
    rep("Gauss", 2),
    round(c(m.mle.gauss.trends[1], m.map.gauss.trends[1]),2),
    round(c(m.mle.gauss.trends[2], m.map.gauss.trends[2]),2),
    round(c(m.mle.gauss.trends[3], m.map.gauss.trends[3]),2),
    round(c(m.mle.gauss.ranges[1], m.map.gauss.ranges[1]),2),
    round(c(m.mle.gauss.ranges[2], m.map.gauss.ranges[2]),2),
    round(c(m.mle.gauss.var, m.map.gauss.var),2)
    )
  )
colnames(estimates_per_estimation) <- c("Estimation", "Kernel", "Trend 1", "Trend 2", "Trend 3", "Range 1", "Range 2", "Variance")
kable(estimates_per_estimation, 
      align=c("l","l","c","c","c","c","c","c"),
      caption = "Estimates table"
      )
```

## Verifying the estimates by-hand for both estimation methods

A good match is found for both the predictions (Figure 10) and the partial derivatives (Figure 11, top row). The two estimation methods produce similar estimates for the partial derivatives (Figure 11, bottom row).

```{r 2nd-functions-work-4, include=F, echo=F}
# make A and A_inv
m.mle.gauss.A <- A_maker(input, m.mle.gauss.ranges)
m.mle.gauss.A.inv <- solve(m.mle.gauss.A)
m.map.gauss.A <- A_maker(input, m.map.gauss.ranges)
m.map.gauss.A.inv <- solve(m.map.gauss.A)

# make t
m.mle.gauss.t <- t_maker(input, testing_input, m.mle.gauss.ranges)
m.map.gauss.t <- t_maker(input, testing_input, m.map.gauss.ranges)
```

```{r, fig.show="hold", out.width="50%", fig.cap="The predictions found by-hand match well with those made by both estimation methods (left), with the latter minus the former found to be essentially zero (right). Top: MLE. Bottom: MAP.", echo=F}

par(mar = c(4, 4, .1, .1))

# MLE plots
m.mle.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.mle.gauss.trends,
                                   m.mle.gauss.t,
                                   m.mle.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.mle.gauss.predict, m.mle.gauss.predictbyhand)

# MAP plots
m.map.gauss.predictbyhand <- byhand_preds(testing_input,
                                   m.map.gauss.trends,
                                   m.map.gauss.t,
                                   m.map.gauss.A.inv,
                                   output,
                                   H)

m_pred_plots_2D(input, output, testing_input, m.map.gauss.predict, m.map.gauss.predictbyhand)

```

```{r 2nd-functions-work-5, include=F, echo=F}
# predictions by hand comparison score
m.mle.gauss.score <- sum(abs(m.mle.gauss.predict$mean - as.vector(m.mle.gauss.predictbyhand))) / (10^(-8)*num_testing_input)
m.map.gauss.score <- sum(abs(m.map.gauss.predict$mean - as.vector(m.map.gauss.predictbyhand))) / (10^(-8)*num_testing_input)

m.mle.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.mle.gauss.pds[,r] <-
    matrix(m.mle.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.mle.gauss.ranges[r]^2)) *
       m.mle.gauss.t) %*% 
    m.mle.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.mle.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.mle.gauss.pdsnormed.2 <- m.mle.gauss.pds / sqrt(rowSums(m.mle.gauss.pds^2))

m.map.gauss.pds <- data.frame(matrix(NA, nrow = num_testing_input, ncol = dim_inputs))
for (r in 1:dim_inputs) {
  m.map.gauss.pds[,r] <-
    matrix(m.map.gauss.trends[r+1], nrow = num_testing_input) + 
    ((2*(matrix(rep(input[,r], num_testing_input), nrow = num_testing_input, byrow = T) - matrix(rep(testing_input[,r], num_obs), nrow = num_testing_input, byrow = F)) / (m.map.gauss.ranges[r]^2)) *
       m.map.gauss.t) %*% 
    m.map.gauss.A.inv %*% (as.matrix(output, ncol = n) - H %*% m.map.gauss.trends)
  # cat("pd_t:", r, "/", dim_inputs, ",", sep = "")
}
m.map.gauss.pdsnormed.2 <- m.map.gauss.pds / sqrt(rowSums(m.map.gauss.pds^2))
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{N}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{N},$$ where $N$ is the number of test points, then divide this by $10^{-8}$ (machine precision) to get `r signif(m.mle.gauss.score, 1)` for MLE and `r signif(m.map.gauss.score, 1)` for MAP.

```{r, fig.show="hold", out.width="50%", fig.cap="Top row: The estimated partial derivatives found by each estimation method (red dots, left MLE and right MAP) match the true partial derivatives (black) well. Bottom plot: The estimated partial derivatives returned by each estimation method are similar (blue for MLE and red for MAP).", echo=F}

par(mar = c(4, 4, .1, .1))

# actual pds
dk_scaled_branin_hoo_expression <- expression((15*x2 - 5/(4 * pi^2) * ((15*x1-5)^2) + 5/pi * (15*x1-5) - 6)^2 + 10 * (1 - 1/(8 * pi)) * cos(15*x1-5) + 10)
# D(dk_scaled_branin_hoo_expression,'x1')
# D(dk_scaled_branin_hoo_expression,'x2')

dk_scaled_branin_hoo_ddx1 <- function(inputs) {
  x1 <- inputs[1]
  x2 <- inputs[2]
  2 * ((5/pi * 15 - 5/(4 * pi^2) * (2 * (15 * (15 * x1 - 5)))) * 
    (15 * x2 - 5/(4 * pi^2) * ((15 * x1 - 5)^2) + 5/pi * (15 * 
        x1 - 5) - 6)) - 10 * (1 - 1/(8 * pi)) * (sin(15 * x1 - 
    5) * 15)
}
dk_scaled_branin_hoo_ddx2 <- function(inputs) {
  x1 <- inputs[1]
  x2 <- inputs[2]
  2 * (15 * (15 * x2 - 5/(4 * pi^2) * ((15 * x1 - 5)^2) + 5/pi * 
    (15 * x1 - 5) - 6))
}
actual.2 <- data.frame(cbind(rep(NA, num_testing_input), rep(NA, num_testing_input)))
colnames(actual.2) <- c("mean1","mean2")
actual.2$mean1 <- matrix(apply(testing_input, 1, dk_scaled_branin_hoo_ddx1), nrow = num_testing_input)
actual.2$mean2 <- matrix(apply(testing_input, 1, dk_scaled_branin_hoo_ddx2), nrow = num_testing_input)

# MLE plots
plot(actual.2, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(actual.2$mean1, m.mle.gauss.pds$X1), 
            max(actual.2$mean1, m.mle.gauss.pds$X1)),
     ylim=c(min(actual.2$mean2, m.mle.gauss.pds$X2), 
            max(actual.2$mean2, m.mle.gauss.pds$X2)),
     pch=19, cex=0.75
     )
points(m.mle.gauss.pds, pch=19, cex=0.75, col="red")
segments(actual.2$mean1, actual.2$mean2,
         m.mle.gauss.pds$X1, m.mle.gauss.pds$X2
         )

# MAP plots
plot(actual.2, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(actual.2$mean1, m.map.gauss.pds$X1), 
            max(actual.2$mean1, m.map.gauss.pds$X1)),
     ylim=c(min(actual.2$mean2, m.map.gauss.pds$X2), 
            max(actual.2$mean2, m.map.gauss.pds$X2)),
     pch=19, cex=0.75
     )
points(m.map.gauss.pds, pch=19, cex=0.75, col="red")
segments(actual.2$mean1, actual.2$mean2,
         m.map.gauss.pds$X1, m.map.gauss.pds$X2
         )

# Estimates comparison plot
plot(m.mle.gauss.pds, xlab="d/dx1", ylab="d/dx2", 
     xlim=c(min(m.map.gauss.pds$X1, m.map.gauss.pds$X1), 
            max(m.map.gauss.pds$X1, m.map.gauss.pds$X1)),
     ylim=c(min(m.map.gauss.pds$X2, m.map.gauss.pds$X2), 
            max(m.map.gauss.pds$X2, m.map.gauss.pds$X2)),
     pch=19, cex=0.75, col="blue"
     )
points(m.map.gauss.pds, pch=19, cex=0.75, col="red")
segments(m.mle.gauss.pds$X1, m.mle.gauss.pds$X2,
         m.map.gauss.pds$X1, m.map.gauss.pds$X2
         )

```

\newpage

# Alignment measure

We calculate the alignment measure of the two functions. As can be see in Figure 12, where the two functions are superimposed, there is a large amount of the 2D input space in which the response surfaces are more orthogonal than they are parallel.

```{r, fig.show="hold", out.width="45%", fig.height = 6, fig.align = "center", echo=F, fig.cap="Contour plots for the function $y = \\sqrt{2 - x_1^2 - x_2^2}$ and the Branin-Hoo function.", warning=F}

par(mar = c(4, 4, .1, .1))

# Data
x <- seq(-2,2, by=0.1)
y <- seq(-2,2, by=0.1)
z_f <- sqrt(2^2 - outer(x ^ 2, y ^ 2, "+"))

contour(x[21:31], y[21:31], z_f[21:31, 21:31], xlim = c(0,1), ylim = c(0,1), nlevels = 40, drawlabels=F, xlab="x1", ylab="x2")

# Data
contour(x.grid,y.grid,z.grid, xlim = c(0,1), ylim = c(0,1), nlevels = 40, drawlabels=F, add = T)
points(input, add=T, pch=19, col="blue")
```

```{r, echo=F}

actual.normed.1 <- actual.1 / sqrt(rowSums(actual.1^2))
actual.normed.2 <- actual.2 / sqrt(rowSums(actual.2^2))

AM.actual.1.2 <- mean(abs(rowSums(actual.normed.1 * actual.normed.2)))
```

```{r, echo=F}
AM.mle.1.2 <- mean(abs(rowSums(m.mle.gauss.pdsnormed.1 * m.map.gauss.pdsnormed.2)))
AM.map.1.2 <- mean(abs(rowSums(m.map.gauss.pdsnormed.1 * m.map.gauss.pdsnormed.2)))
```

Using the **actual** partial derivatives of the two functions at the 100 testing points, we find an AM of 0.6804
<!-- 0.6803849  -->
(and 0.6881 using 1000). 
<!-- 0.6880993  -->
Using MLE at these 100 points we get an alignment for these two functions of 0.6952 (and 0.6937 using 1000)
<!-- 0.6952041 -->
<!-- 0.6937022 -->
and for MAP 0.6900 (and 0.6881 using 1000).
<!-- 0.6900455 -->
<!-- 0.6880806 -->