---
title: "`DiceKriging` vs `RobustGaSP`"
date: "2D test example"
output: pdf_document
fontsize: 10pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{mybgcolor}{HTML}{ebfff0}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RobustGaSP)
library(DiceKriging)
library(knitr)
library(readr)
library(lhs)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html"))
                  )
     ) 
  {
  source(f)
  }
test_function <- "branin"
```

<!-- \pagecolor{mybgcolor} -->

\begin{center}
Using the `r test_function` function
\end{center}

![Fig. 2 from Ensemble of surrogates - Scientific Figure on ResearchGate. Available from: https://www.researchgate.net/figure/Contour-plots-of-two-variable-test-functions-a-Branin-Hoo-function-b-Camelback_fig9_227093096 (accessed 4 Dec 2025). Contour plots of two variable test functions. (a) Branin-Hoo function, (b) Camelback function, (c) Goldstein-Price](U:/ManWin/My Documents/thesis/Rmd/figures_for_Rmd/2D_functions.png)
    
```{r, echo=F}
# dimensional of the inputs
dim_inputs <- 2   
# number of the inputs
num_obs <- 10*dim_inputs
# starting point for data
start_point <- 0
# lhs design
input <- maximinLHS(n=num_obs, k=dim_inputs)  ##maximin lhd sample
## Export 
write_lines(as.vector(input),
            file=
              paste0(
                "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/input_",
                dim_inputs,
                "D_",
                num_obs,
                "obs.txt")
            )

## Import
# input <-
#   matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/input_", dim_inputs, "D_", num_obs, "obs.txt"))),
#          ncol = dim_inputs)

# outputs
output <- matrix(apply(input, 1, test_function), nrow = num_obs)
```

We have a `r dim_inputs`-dimensional input, one output, and `r num_obs` noise-free observations, obtained using the `r test_function` function:

\begin{equation*}
x_1 = \{`r round(input[,1],2)`\}
\end{equation*}
\begin{equation*}
x_2 = \{`r round(input[,2],2)`\}
\end{equation*}
\begin{equation*}
y = \{`r round(output,2)`\}
\end{equation*}

Going forward, we will assume a linear mean function of the form $$m(x) = \beta_0 + \beta_1x_1 + \beta_2x_2.$$ In `RobustGaSP`'s `rgasp()` function, this involves specifying the `trend` matrix $H$:

```{r, echo=FALSE}
H <- H_maker(input)
H
```

as well as setting `zero.mean` equal to `"No"`. The Gauss (squared exponential) *and* Matern 5/2 kernels are compared for both packages, using the same four functions created in **`DiceKriging` vs `RobustGaSP`: 1D test example**.

```{r, echo=F, include=F}
gp_gauss_DK <- gp_km_gauss(input, output)
gp_matern52_DK <- gp_km_matern52(input, output)
```

The output from the two `RobustGaSP` functions is shown below (first for the Gauss kernel and second for the Matern 5/2).

```{r, echo=F}
gp_gauss_RG <- gp_rgasp_gauss(input, output, H)
gp_matern52_RG <- gp_rgasp_matern52(input, output, H)
```

```{r testing-inputs, echo=F}
# number of points to be predicted 
num_testing_input <- 1000

# generate points to be predicted
testing_input <- maximinLHS(n=num_testing_input, k=dim_inputs)

## Export 
# write_lines(as.vector(testing_input),
#             file=
#               paste0(
#                 "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_",
#                 dim_inputs,
#                 "D.txt")
#             )

## Import
testing_input <-
  matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_", dim_inputs, "D.txt"))),
         ncol = dim_inputs)
```

```{r, echo=F}
H_testing <- H_maker(testing_input)
```

```{r perform-predictions, echo=F}
DK_gauss.predict <- gppred_km(gp_gauss_DK)
DK_matern52.predict <- gppred_km(gp_matern52_DK)

RG_gauss.predict <- gppred_rgasp(gp_gauss_RG)
RG_matern52.predict <- gppred_rgasp(gp_matern52_RG)
```

Two checks are carried out. Firstly, the mean width of the 95% CI returned by each of the four set-ups is calculated. Note that I'm getting matching GPs when using `DiceKriging`.

```{r length_of_cred_int, echo=F}
# average length of posterior predictive credible interval

length_95CI_DK_gauss <- length_95CI(DK_gauss.predict$lower95, DK_gauss.predict$upper95)
length_95CI_DK_matern52 <- length_95CI(DK_matern52.predict$lower95, DK_matern52.predict$upper95)
length_95CI_RG_gauss <- length_95CI(RG_gauss.predict@lower95, RG_gauss.predict@upper95)
length_95CI_RG_matern52 <- length_95CI(RG_matern52.predict@lower95, RG_matern52.predict@upper95)

RG_vs_DK_summary <- data.frame(
  cbind(
    c(rep("DiceKriging",2), rep("RobustGaSP",2)),
    rep(c("Gauss", "Matern 5/2"), 2),
    round(c(length_95CI_DK_gauss, length_95CI_DK_matern52, length_95CI_RG_gauss, length_95CI_RG_matern52),4)
    )
  )
colnames(RG_vs_DK_summary) <- c("Package", "Kernel", "Mean width of 95% CI")
kable(RG_vs_DK_summary, 
      align=c("l","l","c"),
      caption = "Mean width of 95% CI"
      )
```

Leave-one-out validation for each of the four set-ups produces the results shown in Tables 2-5 below.

```{r, include=F}
loo_validation_km("gauss")
loo_validation_km("matern52")
loo_validation_rgasp("gauss")
loo_validation_rgasp("matern52")
```

```{r, echo=F}
loo_validation_summary(gp_DK_gauss_loo, "DiceKriging", "Gauss")
```

```{r, echo=F}
loo_validation_summary(gp_DK_matern52_loo, "DiceKriging", "Matern 5/2")
```

```{r, echo=F}
loo_validation_summary(gp_RG_gauss_loo, "RobustGaSP", "Gauss")
```

```{r, echo=F}
loo_validation_summary(gp_RG_matern52_loo, "RobustGaSP", "Matern 5/2")
```

```{r, validation-plots, fig.height=3.5, fig.width=3.5, fig.show="hold", out.width="50%", fig.cap="Xxxxxxx.", echo=F}

par(mar = c(4, 4, .1, .1))

# DiceKriging's plots
loo_validation_plot(gp_DK_gauss_loo)

loo_validation_plot(gp_DK_matern52_loo)

# RobustGaSP's plots
loo_validation_plot(gp_RG_gauss_loo)

loo_validation_plot(gp_RG_matern52_loo)

```

\newpage

Going forward, for now, only the Gauss kernel is dealt with; Matern 5/2 will be returned to once I've figured out the estimates returned by the two packages and the derivative of the Matern 5/2 kernel.

The estimates returned by both packages using the Gauss kernel are shown in Table 6.

```{r, echo=F}
# parameter estimates

gpests_km(gp_gauss_DK, "gauss")
gpests_km(gp_matern52_DK, "matern52")
gpests_rgasp(gp_gauss_RG, "gauss")
gpests_rgasp(gp_matern52_RG, "matern52")

```

```{r, echo=F}
RG_vs_DK_estimates <- data.frame(
  cbind(
    c("DiceKriging", "RobustGaSP"),
    rep("Gauss", 2),
    round(c(gp_dk_gauss_trend_pars[1], gp_rg_gauss_trend_pars[1]),2),
    round(c(gp_dk_gauss_trend_pars[2], gp_rg_gauss_trend_pars[2]),2),
    round(c(gp_dk_gauss_trend_pars[3], gp_rg_gauss_trend_pars[3]),2),
    round(c(gp_dk_gauss_range_pars[1], gp_rg_gauss_range_pars[1] / sqrt(2)),2),
    round(c(gp_dk_gauss_range_pars[2], gp_rg_gauss_range_pars[2] / sqrt(2)),2),
    round(c(gp_dk_gauss_sd_par, gp_rg_gauss_sd_par),2)
    )
  )
colnames(RG_vs_DK_estimates) <- c("Package", "Kernel", "Trend 1", "Trend 2", "Trend 3", "Range 1", "Range 2", "Variance")
kable(RG_vs_DK_estimates, 
      align=c("l","l","c","c","c","c","c","c"),
      caption = "Estimates table"
      )
```

## Verifying the estimates by-hand for both packages

To calculate the posterior predictions at the `r num_testing_input` testing points by-hand, we require the matrix $H$ containing the inputs for the `r num_obs` observations and well as matrix $A$, its inverse, and $\textbf{t}(x^{*})$ for all 1000 of the new inputs $x^{*}$. $A$ and $\textbf{t}(x^{*})$ are built using the estimate of the range parameter returned by each package. 

```{r, echo=F}
# make A and A_inv

# using DiceKriging's range estimate(s)
A_DK <- A_maker_km_gauss(input, gp_dk_gauss_range_pars)
A_inv_DK <- solve(A_DK)

# using RobustGaSP's range estimate(s)
A_RG <- A_maker_rgasp_gauss(input, gp_rg_gauss_range_pars)
A_inv_RG <- solve(A_RG)

# make t

# using DiceKriging's range estimate(s)
t_DK <- t_maker_km_gauss(input, testing_input, gp_dk_gauss_range_pars)

# using RobustGaSP's range estimate(s)
t_RG <- t_maker_rgasp_gauss(input, testing_input, gp_rg_gauss_range_pars)
```

```{r, echo=F}

# DiceKriging's plots
byhand_preds_DK <- byhand_preds(testing_input,
                                   gp_dk_gauss_trend_pars,
                                   t_DK,
                                   A_inv_DK,
                                   output,
                                   H)


# RobustGaSP's plots
byhand_preds_RG <- byhand_preds(testing_input,
                                   gp_rg_gauss_trend_pars,
                                   t_RG,
                                   A_inv_RG,
                                   output,
                                   H)

```

```{r, echo = F}
byhand_scored_DK <- sum(abs(DK_gauss.predict$mean - as.vector(byhand_preds_DK))) / (10^(-8)*num_testing_input)

byhand_scored_RG <- sum(abs(RG_gauss.predict@mean - as.vector(byhand_preds_RG))) / (10^(-8)*num_testing_input)
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{1000}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{1000}$$ then divide this by $10^{-8}$ (machine precision) to get `r signif(byhand_scored_DK, 1)` for `DiceKriging` and  `r signif(byhand_scored_RG, 1)` for `RobustGaSP` (the latter approximately a factor of `r round(byhand_scored_RG / byhand_scored_DK)` larger than the former). By-hand predictions therefore match those returned by `DiceKriging` and `RobustGaSP`.
