---
title: "`RobustGaSP`"
date: "1D test example"
output: pdf_document
fontsize: 10pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{mybgcolor}{HTML}{ffebf5}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RobustGaSP)
library(knitr)
library(readr)
library(lhs)
options(scipen=999)
# source R scripts in
for (f in setdiff(list.files(), 
                  union(union(union(union(union(union(union(list.files(pattern="*.Rmd"), list.files(pattern="*.pdf")),list.files(pattern="*.db")), list.files(pattern="*_files")), list.files(pattern="*.tex")),list.files(pattern="*.log")),list.files(pattern="*.md")),list.files(pattern="*.html"))
                  )
     ) 
  {
  source(f)
  }
```

<!-- \pagecolor{mybgcolor} -->

```{r, echo=F}
# dimensional of the inputs
dim_inputs <- 1    
# number of the inputs
num_obs <- 20
# starting point for data
start_point <- 0
# uniform samples of design
input <- matrix(seq(start_point + 0.1, start_point + 0.9, length.out=num_obs), num_obs)
# Following codes use maximin Latin Hypercube Design, which is typically better than uniform
# library(lhs)
# input <- maximinLHS(n=num_obs, k=dim_inputs)  ##maximin lhd sample

# outputs from 2sin(3x)
test_function_1D <- function(x) {
  2*sin(3*x - 0.5)
  }
output <- matrix(test_function_1D(input), num_obs)

# test_function_1D <- function(x) {
#   2*sin(7*x - 0.5)+0.5
#   }
# output <- matrix(test_function_1D(input), num_obs)
```

We have a `r dim_inputs`-dimensional input, one output, and `r num_obs` noise-free observations, obtained using the function $f(x) = 2\sin(3x - 0.5)$:

\begin{equation*}
x = \{`r input`\}
\end{equation*}
\begin{equation*}
y = \{`r round(output,2)`\}
\end{equation*}

which is plotted in Figure 1.

```{r, echo=F, fig.height=3.5, fig.width=3.5, fig.align="center", fig.cap="1D example observations"}
plot(output ~ input, xlim = c(0,1), pch=19, col = "blue")
```

Going forward, we will assume a linear mean function of the form $$m(x) = \beta_0 + \beta_1x.$$ In `RobustGaSP`'s `rgasp()` function, this involves specifying the `trend` matrix $H$:

```{r, echo=FALSE}
H <- H_maker(input)
H
```

as well as setting `zero.mean` equal to `"No"`. The Gauss (squared exponential) *and* Matern 5/2 kernels are compared, leading to the creating of four functions, `gp_km_gauss()`, `gp_km_matern52()`, `gp_rgasp_gauss()` and `gp_rgasp_matern()`, reprinted below:

```{r, eval=F}
# gp_km_gauss <- function(input_matrix, output_matrix){
#   km(~.,
#      input_matrix,
#      output_matrix,
#      covtype="gauss",
#      optim.method="BFGS",
#      control=list(maxit=500)
#      )
# }

# gp_km_matern52 <- function(input_matrix, output_matrix){
#   km(~.,
#      input_matrix,
#      output_matrix,
#      covtype="matern5_2",
#      optim.method="BFGS",
#      control=list(maxit=500)
#      )
# }

gp_rgasp_gauss <- function(input_matrix, output_matrix, H_matrix){
  rgasp(design = input_matrix,
        response = output_matrix,
        trend = H_matrix,
        lower_bound=FALSE,
        kernel_type = 'pow_exp', alpha = rep(2, nrow(input_matrix)),
        zero.mean = "No"
        )
}

gp_rgasp_matern52 <- function(input_matrix, output_matrix, H_matrix){
  rgasp(design = input_matrix,
        response = output_matrix,
        trend = H_matrix,
        lower_bound=FALSE,
        zero.mean = "No"
  )
}
```

Note that `lower_bound` is set to `FALSE` - Jonathan doesn't *think* this is incorrect.

```{r, echo=F, include=F}
# gp_gauss_DK <- gp_km_gauss(input, output)
# gp_matern52_DK <- gp_km_matern52(input, output)
```

The output from the two `RobustGaSP` functions is shown below (first for the Gauss kernel and second for the Matern 5/2).

```{r, echo=F}
gp_gauss_RG <- gp_rgasp_gauss(input, output, H)
gp_matern52_RG <- gp_rgasp_matern52(input, output, H)
```

We believe there are two optimisations since the first is getting a starting point for the second. Jonathan doesn't think it's a problem that we sometimes get a `FALSE` result for `Convergence`.

```{r testing-inputs, echo=F}
# number of points to be predicted 
num_testing_input <- 1000

# generate points to be predicted
# testing_input <- matrix(sort(runif(num_testing_input*dim_inputs)), num_testing_input,dim_inputs)
amount <- 0.25
testing_input <- matrix(seq(start_point-amount,start_point+1+amount, length.out = num_testing_input))

## Export 
# write_lines(as.vector(testing_input),
#             file=
#               paste0(
#                 "U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_",
#                 dim_inputs,
#                 "D.txt")
#             )

## Import
# testing_input <-
#   matrix(as.numeric(readLines(paste0("U:/ManWin/My Documents/thesis/Rmd/objects_for_Rmd/testing_input_", dim_inputs, "D.txt"))),
#          ncol = dim_inputs)
```

To visualise the posterior mean functions returned by these four functions, predictions are made at `r num_testing_input` new points on the range $[0,1]$. Once again, `rgasp` requires specification of the `trend' matrix $H$, this time containing the testing inputs.

```{r, echo=F}
H_testing <- H_maker(testing_input)
```

```{r perform-predictions, echo=F}
# DK_gauss.predict <- gppred_km(gp_gauss_DK)
# DK_matern52.predict <- gppred_km(gp_matern52_DK)

RG_gauss.predict <- gppred_rgasp(gp_gauss_RG)
RG_matern52.predict <- gppred_rgasp(gp_matern52_RG)
```

These `r num_testing_input` points are plotted in Figure 2 (red line), along with 95% credible intervals (red dashed line) along with the actual function (black line) and `r num_obs` observations (blue points).

```{r Gauss_vs_Matern52, figures-side, fig.show="hold", out.width="50%", fig.cap="Top: DiceKriging with Gauss kernel (left) and Matern 5/2 (right). Bottom: RobustGaSP with Gauss kernel (left) and Matern 5/2 (right).", echo=F}
plot1_function <- function(posterior_mean, posterior_lower95, posterior_upper95){
  plot(testing_input, test_function_1D(testing_input),
       main="",
       xlab="input",
       ylab="output",
       type="l",
       col="black",
       ylim=c(min(posterior_lower95), max(posterior_upper95))
       # ylim=c(-1.1, 2.1)
       )
  lines(testing_input,posterior_mean, col="red")
  lines(testing_input,posterior_lower95, col="red", lty = "dashed")
  lines(testing_input,posterior_upper95, col="red", lty = "dashed")
  points(input, output, pch=19, col="blue")
  legend("topleft",
         c("actual","observations","prediction"),
         fill=c("black","blue","red")
  )
}

par(mar = c(4, 4, .1, .1))
# plot1_function(DK_gauss.predict$mean, DK_gauss.predict$lower95, DK_gauss.predict$upper95)
# plot1_function(DK_matern52.predict$mean, DK_matern52.predict$lower95, DK_matern52.predict$upper95)
plot1_function(RG_gauss.predict@mean, RG_gauss.predict@lower95, RG_gauss.predict@upper95)
plot1_function(RG_matern52.predict@mean, RG_matern52.predict@lower95, RG_matern52.predict@upper95)
```

Various checks are performed (as in the example from `RobustGaSP`'s Help page):

```{r, echo=F}
# actual output values
testing_output <- test_function_1D(testing_input)
```

```{r prop_in_cred_int, echo=F}
# proportion covered by 95% posterior predictive credible interval

# prop_wn_95CI_DK_gauss <- prop_wn_95CI(
#   DK_gauss.predict$lower95, DK_gauss.predict$upper95
#   )
# 
# prop_wn_95CI_DK_matern52 <- prop_wn_95CI(
#   DK_matern52.predict$lower95, DK_matern52.predict$upper95
#   )

prop_wn_95CI_RG_gauss <- prop_wn_95CI(
  RG_gauss.predict@lower95, RG_gauss.predict@upper95
  )

prop_wn_95CI_RG_matern52 <- prop_wn_95CI(
  RG_matern52.predict@lower95, RG_matern52.predict@upper95
  )
```

```{r MSE, echo=F}
# out of sample MSE

# MSE_emulator_DK_gauss <- 
#   MSE(DK_gauss.predict$mean, testing_output)
# 
# MSE_emulator_DK_matern52 <- 
#   MSE(DK_matern52.predict$mean, testing_output)

MSE_emulator_RG_gauss <- 
  MSE(RG_gauss.predict@mean, testing_output)

MSE_emulator_RG_matern52 <- 
  MSE(RG_matern52.predict@mean, testing_output)
```

```{r length_of_cred_int, echo=F}
# average length of posterior predictive credible interval

length_95CI_RG_gauss <- length_95CI(RG_gauss.predict@lower95, RG_gauss.predict@upper95)

length_95CI_RG_matern52 <- length_95CI(RG_matern52.predict@lower95, RG_matern52.predict@upper95)

# length_95CI_DK_gauss <- length_95CI(DK_gauss.predict$lower95, DK_gauss.predict$upper95)
# 
# length_95CI_DK_matern52 <- length_95CI(DK_matern52.predict$lower95, DK_matern52.predict$upper95)
```

```{r checks-summary, echo=F}
RG_vs_DK_summary <- data.frame(
  cbind(
    c(rep("RobustGaSP",2)),
    rep(c("Gauss", "Matern 5/2"), 1),
    round(c(prop_wn_95CI_RG_gauss, prop_wn_95CI_RG_matern52),2),
    signif(c(MSE_emulator_RG_gauss, MSE_emulator_RG_matern52),1),
    round(c(length_95CI_RG_gauss, length_95CI_RG_matern52),4)
    )
  )
colnames(RG_vs_DK_summary) <- c("Package", "Kernel", "Prop. w/n 95% CI", "Out-of-sample MSE", "Mean width of 95% CI")
kable(RG_vs_DK_summary, 
      align=c("l","l","c","c","c"),
      caption = "RobustGaSP manual's suite of checks"
      )
```

Finally, in preparation for the real world data where we cannot perform out-of-sample validation, we'll perform leave-one-out validation: for each of the observations, fitting a GP using all observations except this one, then using the fitted emulator to predict the output at the excluded observation's input setting. The 95% credible interval for each prediction is checked against the actual output value, and the proportion of these CI's that contain the respective observation is found. The results are shown in Tables 2-5 below.

```{r, include=F}
# loo_validation_km("gauss")
# loo_validation_km("matern52")
loo_validation_rgasp("gauss")
loo_validation_rgasp("matern52")
```

```{r, echo=F}
# loo_validation_summary(gp_DK_gauss_loo, "DiceKriging", "Gauss")
```

```{r, echo=F}
# loo_validation_summary(gp_DK_matern52_loo, "DiceKriging", "Matern 5/2")
```

```{r, echo=F}
loo_validation_summary(gp_RG_gauss_loo, "RobustGaSP", "Gauss")
```

```{r, echo=F}
loo_validation_summary(gp_RG_matern52_loo, "RobustGaSP", "Matern 5/2")
```

```{r, validation-plots, fig.show="hold", out.width="50%", fig.cap="LOO cross validation. Top: DiceKriging with Gauss kernel (left) and Matern 5/2 (right). Bottom: RobustGaSP with Gauss kernel (left) and Matern 5/2 (right).", echo=F}

par(mar = c(4, 4, .1, .1))

# DiceKriging's plots
# loo_validation_plot(gp_DK_gauss_loo)
# 
# loo_validation_plot(gp_DK_matern52_loo)

# RobustGaSP's plots
loo_validation_plot(gp_RG_gauss_loo)

loo_validation_plot(gp_RG_matern52_loo)

```

\newpage

Going forward, for now, only the Gauss kernel is dealt with; Matern 5/2 will be returned to once I've figured out the estimates returned by the two packages and the derivative of the Matern 5/2 kernel.

The estimates returned by both packages using the Gauss kernel are shown in Table 2. For `RobustGaSP`, the Manual indicates that it returns "inverse-range parameters", so the inverse of these values have been found, resulting in $\ell_{RK}$. In the Gauss kernel case, the Manual for `RobustGaSP` indicates that the denominator of the fraction is $\ell_{RK}^2$ whereas for `DiceKriging` it's $2\ell_{DK}^2$. To facilitate direct comparison of the parameter estimates, we require a range parameter for `RobustGaSP`, $\ell$, which when plugged into $2\ell^2$ results in $\ell_{RG}^2$. Therefore $$\ell = \frac{\ell_{RG}}{\sqrt{2}}$$ since $$2\left(\frac{\ell_{RG}}{\sqrt{2}}\right)^2 = 2\frac{\ell_{RG}}{2}^2 = \ell_{RG}^2.$$

```{r, echo=F}
# parameter estimates

gpests_km(gp_gauss_DK, "gauss")
gpests_km(gp_matern52_DK, "matern52")
gpests_rgasp(gp_gauss_RG, "gauss")
gpests_rgasp(gp_matern52_RG, "matern52")

```

```{r, echo=F}
RG_vs_DK_estimates <- data.frame(
  cbind(
    c("DiceKriging", "RobustGaSP"),
    rep("Gauss", 2),
    round(c(gp_dk_gauss_trend_pars[1], gp_rg_gauss_trend_pars[1]),2),
    round(c(gp_dk_gauss_trend_pars[2], gp_rg_gauss_trend_pars[2]),2),
    round(c(gp_dk_gauss_range_pars, gp_rg_gauss_range_pars / sqrt(2)),2),
    round(c(gp_dk_gauss_sd_par, gp_rg_gauss_sd_par),2)
    )
  )
colnames(RG_vs_DK_estimates) <- c("Package", "Kernel", "Trend 1", "Trend 2", "Range", "Variance")
kable(RG_vs_DK_estimates, 
      align=c("l","l","c","c","c","c"),
      caption = "Estimates table"
      )
```

## Verifying the estimates by-hand for both packages

To calculate the posterior predictions at the `r num_testing_input` testing points by-hand, we require the matrix $H$ containing the inputs for the `r num_obs` observations and well as matrix $A$, its inverse, and $\textbf{t}(x^{*})$ for all 1000 of the new inputs $x^{*}$. $A$ and $\textbf{t}(x^{*})$ are built using the estimate of the range parameter returned by each package. By-hand predictions match those returned by `DiceKriging` (Figure 2) and `RobustGaSP` (Figure 3).

```{r, echo=F}
# make A and A_inv

# using DiceKriging's range estimate(s)
A_DK <- A_maker_km_gauss(input, gp_dk_gauss_range_pars)
A_inv_DK <- solve(A_DK)

# using RobustGaSP's range estimate(s)
A_RG <- A_maker_rgasp_gauss(input, gp_rg_gauss_range_pars)
A_inv_RG <- solve(A_RG)

# make t

# using DiceKriging's range estimate(s)
t_DK <- t_maker_km_gauss(input, testing_input, gp_dk_gauss_range_pars)

# using RobustGaSP's range estimate(s)
t_RG <- t_maker_rgasp_gauss(input, testing_input, gp_rg_gauss_range_pars)
```

```{r, figures-side, fig.show="hold", out.width="50%", fig.cap="The predictions found by-hand match well with those made by both packages (left), with the latter minus the former found to be essentially zero (right). Top: DiceKriging. Bottom: RobustGaSP.", echo=F}

par(mar = c(4, 4, .1, .1))

# DiceKriging's plots
byhand_preds_DK <- byhand_preds(testing_input,
                                   gp_dk_gauss_trend_pars,
                                   t_DK,
                                   A_inv_DK,
                                   output,
                                   H)

plot_preds_vs_byhand_DK(input, output, testing_input, DK_gauss.predict, byhand_preds_DK)

# RobustGaSP's plots
byhand_preds_RG <- byhand_preds(testing_input,
                                   gp_rg_gauss_trend_pars,
                                   t_RG,
                                   A_inv_RG,
                                   output,
                                   H)

plot_preds_vs_byhand_RG(input, output, testing_input, RG_gauss.predict, byhand_preds_RG)

```

```{r, echo = F}
byhand_scored_DK <- sum(abs(DK_gauss.predict$mean - as.vector(byhand_preds_DK))) / (10^(-8)*num_testing_input)

byhand_scored_RG <- sum(abs(RG_gauss.predict@mean - as.vector(byhand_preds_RG))) / (10^(-8)*num_testing_input)
```

For both we calculate $$\dfrac{\sum\limits_{i=1}^{1000}\Biggl|\text{(predictions from package)}_{i} - \text{(predictions by-hand)}_{i}\Biggr|}{1000}$$ then divide this by $10^{-8}$ (machine precision) to get `r signif(byhand_scored_DK, 1)` for `DiceKriging` and  `r signif(byhand_scored_RG, 1)` for `RobustGaSP` (the latter approximately a factor of `r round(byhand_scored_RG / byhand_scored_DK)` larger than the former).

\textcolor{teal}{Why don't RobustGaSP's estimates change? What is the stochastic element of DiceKriging?}